---
title: "Economic Connectedness"
format: html
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(scales)
library(primer.data)
library(ggplot2)
library(dplyr)
```



The plot below shows the relationship between county population and the total connectedness score, using a simulated dataset. Each point represents a county, and the x-axis is displayed on a logarithmic scale to better visualize differences across a wide range of population sizes. This visualization highlights how connectedness tends to increase with population, though the variation is considerable.



```{r}
#| echo: false
#| warning: false
#| message: false

# Generate county_data with consistent logic
set.seed(123)
county_data <- {
  pop <- 10^runif(100, 3, 6)  # 1,000 to 1,000,000
  data.frame(
    county = paste("County", 1:100),
    population = pop,
    connectedness = rgamma(100, shape = 2, scale = 50) * log10(pop)
  )
}

# First plot
ggplot(county_data, aes(x = population, y = connectedness)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  scale_x_log10(
    labels = trans_format("log10", math_format(10^.x)),
    breaks = trans_breaks("log10", function(x) 10^x)
  ) +
  labs(
    title = "Population Distribution Weighted by Connectedness",
    subtitle = "Counties with larger populations tend to have higher total connectedness scores.",
    y = "Sum of Connectedness Scores",
    x = "Population (log scale)",
    caption = "Data source: simulated data\nNote: Populations shown on logarithmic scale"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    plot.caption = element_text(size = 8, hjust = 0)
  )
```


This plot presents the same relationship between population and connectedness but applies a smoothing curve (LOESS) to highlight the overall trend. The x-axis is again shown on a log scale, but using a different transformation function to support smoothing. Counties with missing data are excluded. The LOESS line helps reveal that connectedness increases non-linearly with population, especially in mid-to-large sized counties.


```{r}
#| echo: false
#| warning: false
#| message: false

# Clean data by removing missing values
connectedness_clean <- county_data %>%
  drop_na(population, connectedness)

# Informational message
message(paste("Removed", nrow(county_data) - nrow(connectedness_clean), "rows with missing values"))

# Second plot
ggplot(connectedness_clean, aes(x = population, y = connectedness)) +
  geom_point(alpha = 0.5, color = "#2c7fb8", size = 2) +
  scale_x_continuous(
    trans = "log10",
    labels = comma,
    breaks = 10^(3:6)
  ) +
  geom_smooth(
    method = "loess",
    color = "black",
    se = FALSE,
    linewidth = 1
  ) +
  labs(
    title = "County Connectedness by Population",
    subtitle = "Larger populations tend to have higher connectedness scores",
    x = "Population (log scale)",
    y = "Connectedness Score",
    caption = paste("Data from simulated county data |",
                    nrow(county_data) - nrow(connectedness_clean),
                    "counties with missing data excluded")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.minor = element_blank()
  )
```